---
title: "R Visualization"
author: "Kyle Chung, Ben Chen, Mansun Kuo"
date: '`r Sys.Date()`'
output:
  ioslides_presentation:
    fig_height: 4
    fig_width: 7
    widescreen: yes
subtitle: Using ggplot2
---

```{r include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = TRUE,
                      echo = TRUE,
                      message = TRUE,
                      fig.align='center',
                      cache=TRUE)
```



```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(gridExtra)
library(dplyr)
```

# Bar Plot

# Line Graph

# Scatter Plot
## IMDB data
```{r}
movies1 <- movies[!is.na(movies$budget),]
ggplot(movies1, aes(x=budget, y=rating)) + geom_point()
```


## Control the shape & size of points
```{r}
ggplot(movies1, aes(x=budget, y=rating)) + geom_point(shape=5, size=3)
```


## All point shape types in `ggplot2`
```{r,fig.align='center', fig.retina=2, fig.height=7, fig.width=7, echo=FALSE}
spoint <- data.frame(x=1,y=1)
symbol_points <- list()
for (i in 1:25) {
    symbol_points[[i]] <- ggplot(spoint, aes(x=x,y=y)) + 
        geom_point(shape=i, size=5) +
        ggtitle(sprintf('shape=%s',i)) + 
        theme(axis.text.x=element_blank(), axis.text.y=element_blank()) +
        xlab('') + ylab('')
}
symbols <- do.call(arrangeGrob,symbol_points)
symbols
```


## Grouping: by binary variable

This usually happens accidentally.

```{r}
ggplot(movies1, aes(x=budget, y=rating, color=Action)) + geom_point()
```

    
## Grouping: by categarical variable
```{r}
ggplot(movies1, aes(x=budget, y=rating, color=factor(Action))) + 
    geom_point() + labs(color='Action Movie?')
```

    
## Multi-grouping
```{r}
ggplot(movies1, aes(x=budget, y=rating, color=factor(Action), shape=(length > 120))) + 
    geom_point(size=3) + labs(color='Action Movie?')
```


## Fit regression line
```{r}
ggplot(movies, aes(x=votes, y=rating)) + geom_point() +
    stat_smooth(method=lm, level=.95) # add se=FALSE to disable CI
```


## The default is a polynomial fit
```{r message=FALSE}
ggplot(movies, aes(x=votes, y=rating)) + geom_point() + stat_smooth()
```


## Fitting is smart to align with grouping
```{r}
ggplot(movies1, aes(x=budget, y=rating, color=factor(Action))) + 
    geom_point() + labs(color='Action Movie?') + stat_smooth(method=lm, se=FALSE)
```


## What if the model is pre-computed?
```{r}
lm_model <- lm(rating ~ budget, data=movies1)
ggplot(movies1, aes(x=budget, y=rating)) + geom_point() +
    geom_line(aes(x=budget, y=lm_model$fitted.values), color='blue')
```


## Scatter plot "as is": Using `geom_text`
```{r}
starmovies <- movies[movies$votes > mean(movies$votes),]
starmovies <- starmovies[order(-starmovies$rating),][1:10,]
ggplot(starmovies, aes(x=votes, y=rating)) + geom_point() + geom_text(aes(label=title))
```


## Fine-tune
```{r}
ggplot(starmovies, aes(x=votes, y=rating)) + geom_point(color='red') + 
    geom_text(aes(label=title), hjust=0, vjust=0, angle=20) +
    xlim(0, max(starmovies$votes)*2) +
    ylim(min(starmovies$rating), 9.2)
```


## The Final Challenge:

Which type of film cost the most, averagely?

```{r}
movietype <- colnames(movies)[18:24]
movies1_singletype <- movies1[rowSums(movies1[, movietype]) == 1,] # remove multi-typed
# mean(movies1_singletype[movies1_singletype$Animation == 1, 'budget'])
lmfit <- lm(as.formula(paste('budget ~', paste(movietype, collapse=' + '))), 
            movies1_singletype)
summary(lmfit)$coef # where is type 'Short'?
```


## Try to plot this!

What is the association between cost and rating, conditional on type?

```{r echo=FALSE}
dummies <- as.matrix(movies1_singletype[, movietype])
movies1_singletype$Type <- factor(dummies %*% (1:length(movietype)), labels=movietype) 
ggplot(movies1_singletype, aes(x=budget, y=rating, color=Type)) + 
  geom_point() +
  stat_smooth(method=lm, se=FALSE, fullrange=FALSE, size=1.5) +
  scale_color_discrete(name='Movie Type: # of samples', labels=paste(levels(movies1_singletype$Type), ': ', table(movies1_singletype$Type)))
```


## A Hint:
You need to convert multiple dummies into one factor

+ Before:
```{r echo=FALSE}
dd <- as.matrix(movies1_singletype[, movietype])
head(dd)
```
+ After:
```{r echo=FALSE}
head(factor(dd %*% (1:length(movietype)), labels=movietype))
```


## Exercise: Answer
```{r eval=FALSE}
# convert multiple dummies into one factor as grouping var
# a little matrix operation will do the trick
dummies <- as.matrix(movies1_singletype[, movietype])
movies1_singletype$Type <- factor(dummies %*% (1:length(movietype)), labels=movietype)

# The rest of the task is rather simple
ggplot(movies1_singletype, aes(x=budget, y=rating, color=Type)) + 
  geom_point() +
  
  # set fullrange=T will extend the fitted line outside the sample range
  stat_smooth(method=lm, se=FALSE, fullrange=FALSE, size=1.5) +
  
  # color is the grouping interface, hence scale_color_*
  scale_color_discrete(name='Movie Type: # of samples', 
                       labels=paste(levels(movies1_singletype$Type), ': ', 
                                    table(movies1_singletype$Type)))
```


## The regression problem behind the scene
```{r}
interact_terms <- paste(paste(movietype, '*budget', sep=''), collapse=' + ')
lmfit <- lm(as.formula(paste('rating ~', interact_terms)), movies1_singletype)
tail(summary(lmfit)$coef)
```
+ None of the interactive term is statistically significant, indeed
+ **Visualization != Analysis** (Our eyes were not born to work on numbers.)
+ Plots can be easily manipluated to be misleading, accidentally or **on purpose**


## References
+ [ggplot2 Official document](http://ggplot2.org/)
+ [R Graphics Cookbook](http://www.cookbook-r.com/Graphs/)
+ [Source code of this slide](https://github.com/everdark/lecture_ggplot)
    + The source is tested only on OS X 10.9.3
+ [Introduction to Programming R](https://github.com/everdark/lecture_rintro)
    + Knowlege of general programming is requisite

# Bonus: Annotation


## Annotation
```{r}
plot(movies1$budget, movies1$rating) # base solution
abline(h=median(movies1$rating), col='red')
text(x=max(movies1$budget)*.9, y=median(movies1$rating), 
     labels='Median of Rating', col='red', pos=1)
```


## Annotation: Add lines
```{r fig.show='hold',fig.height=3.5}
brggp <- ggplot(movies1, aes(x=budget, y=rating)) + geom_point() 
brggp + geom_hline(yintercept=median(movies1$rating)) # ?geom_abline for general setup
# brggp + geom_hline(data=movies1, aes(yintercept=median(rating)))  # the same
# brggp + geom_hline(aes(yintercept=median(movies1$rating)))        # the same
```


## Annotation: Add (single) texts
```{r,fig.show='hold',fig.height=3.5}
brggp + geom_hline(yintercept=median(movies1$rating), color='red') + 
  annotate('text', x=Inf, y=median(movies1$rating), 
           label='Medaion of Rating', color='red', vjust=1.2, hjust=1)
# Don't use geom_text for single annotation to avoid overplotting
```


## Annotation: Add segments
```{r,fig.height=3.5}
shaw <- movies1[grep('Shawshank Redemption', movies1$title, fixed=TRUE),]
brggp + annotate('segment', xend=shaw$budget, yend=shaw$rating, x=Inf, y=-Inf,
                 arrow=grid::arrow(), color='red') +
  annotate('text', label='The Shawshank Redemption?', x=Inf, y=-Inf,
           hjust=1.5, vjust=-1, color='red')
```


## Annotation: Add shaded area
```{r,fig.height=3.5}
yearcount <- aggregate(title ~ year, data=movies, FUN=length)
ggplot(yearcount, aes(x=year, y=title)) + geom_line() +
  annotate('rect', xmin=1990, xmax=2000, ymin=-Inf, ymax=Inf, fill='blue', alpha=.25)
```

# Bonus: Facet / Multi-plotting


## Facet: Single grouping
```{r,message=FALSE}
gg <- ggplot(movies1_singletype, aes(x=rating, y=..density..)) + geom_bar()
gg + facet_grid(Action ~ .) # Plot with grouping variable in different window (Vertical)
```


## Facet: Single grouping
```{r,message=FALSE, results='hold'}
# Plot with grouping variable in different window (Horizontal)
gg + facet_grid(. ~ Action) 
```


## Facet: Multiple grouping
```{r,message=FALSE}
movies1_singletype$modern <- (movies1_singletype$year > 2000)
ggplot(movies1_singletype, aes(x=rating, y=..density..)) + 
  geom_bar() + facet_grid(modern ~ Action)
```


## Facet: Multi-layer grouping
```{r,message=FALSE}
movies1_singletype$rated <- (movies1_singletype$mpaa != '')
ggplot(movies1_singletype, aes(x=rating, color=modern)) + 
  geom_line(stat="density") + facet_grid(Type ~ rated)
```


## Facet: Change labels
+ Way 1: Change the grouping var in data.frame to reflect the alternate labels
+ Way 2: Write customized labeller function for `facet_grid`
```{r,message=FALSE}
print(label_value) # the default labeller plugg in facet_grid(..., labeller)
mylabeller <- function(variable, value){
  if ( variable=='rated' ) 
    value <- ifelse(value == TRUE, 'Rated Movies', 'Unrated Movies')
  else if ( variable=='Type' ) 
    as.character(value)
}
```


## Facet: Change labels
```{r,message=FALSE}
ggplot(movies1_singletype, aes(x=rating, color=modern)) +
  geom_line(stat="density") + facet_grid(Type ~ rated, labeller=mylabeller)
```


## Multi-plotting by `gridExtra` (1/3)

```{r,fig.align='center', fig.retina=2, fig.height=2, fig.width=2}
library(gridExtra)
drawPoint <- function(i) {
  ggplot(data.frame(x=1, y=1), aes(x=x,y=y)) + 
    geom_point(shape=i, size=5) +
    ggtitle(sprintf('shape=%s',i)) + 
    theme(axis.text.x=element_blank(), axis.text.y=element_blank()) +
    xlab('') + ylab('')
  }
drawPoint(25)
```


## Multi-plotting by `gridExtra` (2/3)

```{r,fig.align='center', fig.retina=2, fig.height=5.5, fig.width=5.5, fig.show='hide'}
symbol_points <- mapply(drawPoint, 1:25, SIMPLIFY=FALSE)
symbols <- do.call(arrangeGrob, symbol_points)
symbols
```


## Multi-plotting by `gridExtra` (3/3)

```{r,fig.align='center', fig.retina=2, fig.height=5.5, fig.width=5.5, echo=FALSE}
symbols
```

